Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-22 paper.qmd
Added plot for MSE Vs bin counts for different lwd thresholds

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve,
    df_bin = df_bin_scurve,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(bin1 = xbins,
           bin2 = bin2,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve),
           a1 = a1,
           mean_counts = sum(df_bin_centroids_scurve$bin_counts)/NROW(df_bin_centroids_scurve))

@@

count_mse_scurve <- ggplot(error_scurve,
                     aes(x = mean_counts,
                         y = MSE)) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) +
  # geom_vline(xintercept = 0.116,
  #          linetype=2,
  #          color = "#bdbdbd",
  #          linewidth=1) +
  # geom_hline(yintercept = 52,
  #          linetype=2,
  #          color = "#bdbdbd",
  #          linewidth=1) +
  # geom_vline(xintercept = 11, linetype="solid",
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 9, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 13, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  labs(x = expression(mean(n[h])), y = "MSE") +
  ggtitle("(c)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 10,
                                   angle = 90),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12,
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))

@@

```{r}
#| label: distribution-scurve
## To initialize effective bins along x
effective_bin1_scurve <- 14

@@

```{r}
#| label: bin-counts-scurve

@@

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve_high_dens,
    df_bin = df_bin_scurve_high_dens,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(benchmark_rm_lwd = benchmark_rm_lwd,
           bin1 = effective_bin1_scurve,
           bin2 = effective_bin2_scurve,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve_high_dens))

@@

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = benchmark_rm_lwd, 
                         y = MSE)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.3
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  scale_x_continuous("standardized bin count", 
         transform = "reverse") + 
  ylab("MSE") +
  ggtitle("(c)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        plot.title = element_text(size = 12))

@@

mse_scurve_b + a1_m_scurve +
  count_mse_scurve + distance_hist + 
  plot_layout(ncol=2)
```

>>>>

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve,
    df_bin = df_bin_scurve,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(bin1 = xbins,
           bin2 = bin2,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve),
           a1 = a1)

@@

@@

```{r}
#| label: distribution-scurve
## To initialize effective bins along x
effective_bin1_scurve <- 11

@@

```{r}
#| label: bin-counts-scurve
#| eval: false

@@

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve_high_dens,
    df_bin = df_bin_scurve_high_dens,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(benchmark_rm_lwd = benchmark_rm_lwd,
           bin1 = effective_bin1_scurve,
           bin2 = effective_bin2_scurve,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))

@@

error_rm_scurve <- error_rm_scurve |>
  mutate(mean_counts = round(mean_counts, 0)) |>
  filter(b_non_empty >= 5)

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = mean_counts, 
                         y = MSE)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.3
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = unique(error_rm_scurve$mean_counts)) +
  xlab(expression(mean(n[h]))) +
  ylab("MSE") +
  ggtitle("(c)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        plot.title = element_text(size = 12),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5))

@@

mse_scurve_b + a1_m_scurve +
  mse_scurve_lwd + distance_hist + 
  plot_layout(ncol=2)
```
